---
title: "SharkDistribution"
author: "Max Schwarz"
date: "12 July 2019"
output: html_document
---

```{r load_libraries, include = FALSE}
library(rgbif)
library(tidyverse)
library(maps)
library(leaflet)
library(sf)
library(rgdal)
library(ggmap)
library(maptools)
library(raster)
#library(devtools)
#library(repurrrsive)
```

---
# Mapping Endangered Shark Species Distribution \
\

## 1) Downloading Data from GBIF

<!-- Will explain what I have done in this text chunk -->

```{r dowload_data, include = FALSE}
source("functions.R") #call functions from here

#create vectors (sequence of data elements of same type) for all orders

#Carcharhiniformes

carc <- c('Mustelus fasciatus', 'Cephaloscyllium albipinnum', 'Isogomphodon oxyrhynchus', 'Glyphis garricki', 'Carcharhinus hemiodon', 'Glyphis gangeticus', 'Haploblepharus kistnasamyi', 'Sphyrna mokarran', 'Sphyrna lewini', 'Lamiopsis temminckii', 'Carcharhinus leiodon', 'Holohalaelurus punctatus', 'Mustelus schmitti', 'Holohalaelurus favus', 'Carcharhinus borneensis', 'Eusphyra blochii', 'Carcharhinus dussumieri', 'Glyphis glyphis', 'Hemitriakis leucoperiptera')

#Orectolobiformes

orec <- c('Rhincodon typus', 'Stegostoma fasciatum')

#Squatiniformes

squat <- c('Squatina guggenheim', 'Squatina squatina', 'Squatina formosa', 'Squatina oculata', 'Squatina occulta', 'Squatina argentina', 'Squatina aculeata')

#Squaliformes

squal <- c('Squalus chloroculus', 'Centrophorus harrissoni')

#Lamniformes

lamn <- c('Isurus paucus', 'Isurus oxyrinchus')

#create a vector composed of all orders

species <- c(carc, orec, squat, squal, lamn)

#create a list called distributions, which will house all species data

distributions <- list()
 
for(sp in species) {
  distributions[[sp]] <- import_data(sp)
}

#saving all data in distributions dataframe

saveRDS(distributions, file = "Data/processed/shark_data.rds")
```


## 2) Reading in MPA shapefiles

<!-- Will explain what I have done in this text chunk -->

```{r import_shapefile, include = FALSE}
#reading in the global MPA shapefile

GMPA <- sf::read_sf(dsn = "Data/WDPA July 2019 MPAs", layer = "WDPA_Jul2019_marine-shapefile-polygons")
GMPA2 <- GMPA %>% ggplot2::fortify()

saveRDS(GMPA2, "Data/processed/gmpa2.rds")

GMPA2 <- readRDS("Data/processed/gmpa2.rds")
```

## Challenge 1: Subset MPA object without calling filter on the whole thing, because this crashes Max's computer

To do this create a version without geometries. We can use that to find the row numbers we want to extract

```{r}
GMPA2_index <- readRDS("Data/processed/gmpa2_marine_only.rds")
GMPA2_index$geometry <- NULL
saveRDS(GMPA2_index, "Data/processed/gmpa2_index.rds")
```

use the small index to more efficiently sample the GMPA datasets:

```{r}
GMPA2_index %>% filter(PARENT_ISO == "AUS")
```

we want the row numbers for these:

```{r}
i <- which(GMPA2_index$PARENT_ISO == "AUS") 

AUS_MPAs <- GMPA2[i, ]
```

```{r}
library(leaflet)

leaflet() %>% 
  addTiles() %>% 
  addPolygons(data=AUS_MPAs, color = "grey", fillColor = "orange")
```

```{r}
leaflet() %>% 
  addTiles() %>% 
  addPolygons(data=one_mpa, color = "grey", fillColor = "orange")
```

## Challenge 2: Get centroids of all MPAS

```{r}
GMPA2 <- readRDS("Data/processed/gmpa2_marine_only.rds")

GMPA2_centroid <- st_centroid(GMPA2)
saveRDS(GMPA2_centroid, "Data/processed/gmpa2_centroid.rds")

GMPA2_centroid <- readRDS("Data/processed/gmpa2_centroid.rds")

```


Check by plotting:
```{r}
world <- map_data("world")
ggplot() + 
  geom_map(data=world, map=world,
                    aes(x=long, y=lat, map_id=region),
                    color="white", fill="#7f7f7f", size=0.05, alpha=1/4) +
  geom_sf(data = GMPA_centroid, aes(colour="blue"))
 
```



## 3) Creating distribution maps

Load data

```{r}
source("functions.R")
distributions <- readRDS("Data/processed/shark_data.rds")
distributions_all <- distributions %>% bind_rows()
GMPA2 <- readRDS("Data/processed/gmpa2.rds")
```

Clean up data

```{r}
distributions[[1]] <- distributions[["Mustelus fasciatus"]][-c(1:2),]
distributions[[2]] <- distributions[["Cephaloscyllium albipinnum"]][-c(74, 76),]
distributions[[4]] <- distributions[["Glyphis garricki"]][-c(17:19, 21),]

min_lat <- min(distributions[[4]][,3])
max_lat <- max(distributions[[4]][,3])
min_lng <- min(distributions[[4]][,4])
max_lng <- max(distributions[[4]][,4])

saveRDS(distributions, "Data/processed/updated_shark_data.rds")
```

Interactive using leaflet:

```{r}
source("functions.R")
create_leaflet_map("Mustelus fasciatus", distributions)
create_leaflet_map("Cephaloscyllium albipinnum", distributions)
create_leaflet_map("Isogomphodon oxyrhynchus", distributions)
create_leaflet_map("Glyphis garricki", distributions)
```

All species
```{r}

distributions %>% bind_rows() %>% 
    leaflet() %>%
    addTiles() %>%
    addCircleMarkers(lat = ~latitude, lng = ~longitude, radius = 6, stroke = FALSE, 
                     label = ~species_name, fillOpacity = 0.5)

#(how to colour by species?)
```


Static maps using ggplot

```{r}
brazil_region <- borders("world", region = c("Brazil", "Uruguay", "Argentina"), fill = "grey70", colour = "black")

ggplot() + brazil_region + theme_bw() + xlab("Longitude") + ylab("Latitude") + 
  theme(panel.border = element_blank(), panel.grid.major = element_line(colour = "grey80"), panel.grid.minor = element_blank()) +
       geom_point(data = distributions[["Mustelus fasciatus"]], aes(x = longitude, y = latitude))


my_map <- 
  ggplot() + geom_map(data=world, map=world,
                    aes(x=long, y=lat, map_id=region),
                    color="white", fill="#7f7f7f", size=0.05) +
#       geom_sf(data = GMPA2 %>% slice(1:10), aes(colour="orange")) +
       geom_point(data = distributions[["Mustelus fasciatus"]], aes(x = longitude, y = latitude))
my_map
```

Plot of an MPA:
```{r}
ggplot() + 
  geom_sf(data = URY_MPAs, aes(fill = "tomato", colour ="gray"))

basemap2 <- get_map(location = c(left = -57, bottom = -35, right = -53, top = -32.5), zoom = 10, maptype = "roadmap", source = "google", force = TRUE)

ggmap(basemap2) +
  geom_sf(data = df, aes(fill = "tomato", colour ="gray"))

URY_MPAs <- readRDS("Data/MPAs by country/URY_MPAs.rds")

test1 <- URY_MPAs %>% purrr::map(function(df)rename(df, x = latitude, y = longitude))

glimpse(URY_MPAs$geometry[[7]])

mylist <- URY_MPAs$geometry
mylist2 <- mylist[[1]]
mylist3 <- mylist2[[1]]
mylist4 <- mylist3[[1]]

df <- data.frame(mylist4)

colnames(mylist4) <- c("lon", "lat")

colnames(mylist[[8]][[1]][[1]]) <- c("lon", "lat")

colnames(URY_MPAs$geometry[[1]][[1]][[1]]) <- c("lon", "lat")

#NEED TO KNOW WHAT COLUMN NAMES FOR LONG AND LAT ARE
  
```

Plot of two MPAs:

```{r}
ggplot() + geom_sf(data = GMPA2 %>% slice(1:2), aes(colour="orange"))

```


A function for plotting distributions:

```{r}
plot_species_distn <- function(species_name, distributions) {
world <- map_data("world")
  ggplot() + geom_map(data=world, map=world,
                    aes(x=long, y=lat, map_id=region),
                    color="white", fill="#7f7f7f", size=0.05, alpha=1/4) +
       geom_point(data = distributions[[species_name]], aes(x = longitude, y = latitude))
     }

plot_species_distn("Cephaloscyllium albipinnum", distributions)
plot_species_distn('Mustelus fasciatus', distributions)


ggplot() + 
  geom_map(data=world, map=world,
                    aes(x=long, y=lat, map_id=region),
                    color="white", fill="#7f7f7f", size=0.05, alpha=1/4) +
  geom_point(data = distributions_all, aes(x = longitude, y = latitude)) + 
  facet_wrap(~species_name)
  
```

Using leaflet on polygons:

```{r}

leaflet(GMPA2 %>% slice(1:100)) %>% 
  addTiles() %>% 
  addPolygons()

```

Areas of mpas

```{r}
# calculate areas
areas <- st_area(GMPA2)
```

Convert distributions to sf objects:
```{r}
distributions_all_sf <- st_as_sf(distributions[[1]], coords = c("longitude", "latitude"), 
                 crs = 4326,  agr = "constant")
```

Polygon centres

```{r}
# Calculate centroids
GMPA2_c <- st_centroid(GMPA2)

world <- map_data("world")
ggplot() + 
  geom_map(data=world, map=world,
                    aes(x=long, y=lat, map_id=region),
                    color="white", fill="#7f7f7f", size=0.05, alpha=1/4) +
  geom_sf(data = GMPA2_c, aes(colour="blue")) +
  geom_point(data = distributions_all_sf, aes(col="organge")) 
 
```

Crop to particular area (Australia)

```{r}
GMPA2_aus <- st_crop(GMPA2, xmin = 100, ymin = -50, xmax = 150, ymax = -25)
distributions_aus <- st_crop(distributions_all_sf, xmin = 100, ymin = -50, xmax = 150, ymax = -25)

# centre of mpa

GMPA2_aus_c <- st_centroid(GMPA2_aus)

```

```{r}

ggplot(GMPA2_aus_c) + 
  geom_sf() + 
  geom_sf(data = distributions_aus, aes(col="organge"))

```

Some plots with Mapview

```{r}
library(mapview)
mapview(distributions_all_sf, layer.name = "Mus") + mapview(Mus_MPAs, layer.name= "MPAs", color="orange")

```

Check distance between each point and centroid of closest MPAs

```{r}
nearest_mpa <- st_nearest_points(distributions_aus, GMPA2_aus_c)
distances <- st_length(nearest_mpa)
```

Make subsets of GMPA2 data

```{r}
# Selecting only important column names

GMPA2 <- GMPA2 %>% dplyr::select(NAME, DESIG_ENG, MARINE, REP_M_AREA, STATUS, STATUS_YR, PARENT_ISO)

# Looking for data entries with MARINE value 0 -> 0 = 100% terrestrial PA

PA <- GMPA2 %>% dplyr::select(MARINE)
PA$geometry <- NULL
PA <- PA %>% dplyr::count(MARINE)

# Removing values of 0 in MARINE column (0 = 100% terrestrial PA)

GMPA2 <- subset(GMPA2, MARINE != 0)
saveRDS(GMPA2, "Data/processed/gmpa2_marine_only.rds")
GMPA2 <- readRDS("Data/processed/gmpa2_marine_only.rds")

# Creating subsets of data - to make things more manageable

tempGMPA2 <- head(GMPA2, n = 100)

sub01 <- GMPA2[1:1000,]
sub02 <- GMPA2[1001:2000,]
sub03 <- GMPA2[2001:3000,]
sub04 <- GMPA2[3001:4000,]
sub05 <- GMPA2[4001:5000,]
sub06 <- GMPA2[5001:6000,]
sub07 <- GMPA2[6001:7000,]
sub08 <- GMPA2[7001:8000,]
sub09 <- GMPA2[8001:9000,]
sub10 <- GMPA2[9001:10000,]
sub11 <- GMPA2[10001:11000,]
sub12 <- GMPA2[11001:12000,]
sub13 <- GMPA2[12001:13000,]
sub14 <- GMPA2[13001:14000,]
sub15 <- GMPA2[14001:14400,]

# Counting countries - useful for country specific subsets

countries_name <- GMPA2 %>% dplyr::select(PARENT_ISO)
countries_name$geometry <- NULL
countries_name <- countries_name %>% dplyr::count(PARENT_ISO)
saveRDS(countries_name, "Data/processed/countries.rds")
countries_name <- readRDS("Data/processed/countries.rds")

# Creating subsets of data corresponding to each country (PARENT_ISO)

ABNJ_MPAs <- sub01 %>% filter(PARENT_ISO == "ABNJ")
saveRDS(ABNJ_MPAs, file = "Data/MPAs by country/ABNJ_MPAs.rds")

# Mustelus fasciatus regions

ARG_MPAs1 <- sub09 %>% filter(PARENT_ISO == "ARG")
ARG_MPAs2 <- sub10 %>% filter(PARENT_ISO == "ARG")
ARG_MPAs <- rbind(ARG_MPAs1, ARG_MPAs2)
saveRDS(ARG_MPAs, file = "Data/MPAs by country/ARG_MPAs.rds")

BRA_MPAs1 <- sub02 %>% filter(PARENT_ISO == "BRA")
BRA_MPAs2 <- sub09 %>% filter(PARENT_ISO == "BRA")
BRA_MPAs3 <- sub10 %>% filter(PARENT_ISO == "BRA")
BRA_MPAs <- rbind(BRA_MPAs1, BRA_MPAs2, BRA_MPAs3)
saveRDS(BRA_MPAs, file = "Data/MPAs by country/BRA_MPAs.rds")

URY_MPAs <- sub05 %>% filter(PARENT_ISO == "URY")
saveRDS(URY_MPAs, file = "Data/MPAs by country/URY_MPAs.rds")

ARG_MPAs <- readRDS("Data/MPAs by country/ARG_MPAs.rds")
BRA_MPAs <- readRDS("Data/MPAs by country/BRA_MPAs.rds")
URY_MPAs <- readRDS("Data/MPAs by country/URY_MPAs.rds")
Mus_MPAs <- rbind(ARG_MPAs, BRA_MPAs, URY_MPAs)
saveRDS(Mus_MPAs, file = "Data/Final MPAs/Mus_MPAs.rds")

# Cephaloscyllium albipinnum region

AUS_MPAs1 <- sub01 %>% filter(PARENT_ISO == "AUS")
AUS_MPAs2 <- sub02 %>% filter(PARENT_ISO == "AUS")
AUS_MPAs3 <- sub03 %>% filter(PARENT_ISO == "AUS")
AUS_MPAs4 <- sub04 %>% filter(PARENT_ISO == "AUS")
AUS_MPAs <- rbind(AUS_MPAs1, AUS_MPAs2, AUS_MPAs3, AUS_MPAs4)
saveRDS(AUS_MPAs, file = "Data/MPAs by country/AUS_MPAs.rds")
saveRDS(AUS_MPAs, file = "Data/Final MPAs/Cep_MPAs.rds")

# Isogomphodon oxyrhynchus region

SUR_MPAs <- sub04 %>% filter(PARENT_ISO == "SUR")
saveRDS(SUR_MPAs, "Data/MPAs by country/SUR_MPAs.rds")

TTO_MPAs <- sub04 %>% filter(PARENT_ISO == "TTO")
saveRDS(TTO_MPAs, "Data/MPAs by country/TTO_MPAs.rds")

VEN_MPAs <- sub06 %>% filter(PARENT_ISO == "VEN")
saveRDS(VEN_MPAs, "Data/MPAs by country/VEN_MPAs.rds")

Iso_MPAs <- rbind(SUR_MPAs, TTO_MPAs, VEN_MPAs)
saveRDS(Iso_MPAs, "Data/Final MPAs/Iso_MPAs.rds")

# Glyphis garricki region (RIVER SHARK)

# Uses AUS_MPAs

PNG_MPAs <- sub04 %>% filter(PARENT_ISO == "PNG")
saveRDS(PNG_MPAs, "Data/MPAs by country/PNG_MPAs.rds")

Gly_MPAs <- rbind(AUS_MPAs, PNG_MPAs)
saveRDS(Gly_MPAs, "Data/Final MPAs/Gly_MPAs.rds")

CHL_MPAs1 <- sub10 %>% filter(PARENT_ISO == "CHL")
CHL_MPAs2 <- sub15 %>% filter(PARENT_ISO == "CHL")
CHL_MPAs <- rbind(CHL_MPAs1, CHL_MPAs2)
saveRDS(CHL_MPAs, file = "Data/MPAs by country/CHL_MPAs.rds")

FIN_MPAs <- sub01 %>% filter(PARENT_ISO == "FIN")

NZL_MPAs <- sub01 %>% filter(PARENT_ISO == "NZL")
```

Create maps of specific regions

```{r}
library(rmapshaper)

CHL_MPAs <- readRDS("Data/MPAs by country/CHL_MPAs.rds")
leaflet(CHL_MPAs) %>% 
  addTiles() %>% 
  addPolygons()

ABNJ_MPAs <- readRDS("Data/MPAs by country/ABNJ_MPAs.rds")
simplified <- rmapshaper::ms_simplify(ABNJ_MPAs)
leaflet() %>% 
  addTiles() %>% 
  addPolygons(data = simplified, color = "#444444", weight = 1, smoothFactor = 0.5,
    opacity = 1.0, fillOpacity = 0.5, fillColor = "gray30") %>%
  addCircleMarkers(data = distributions[["Mustelus fasciatus"]], lat = ~latitude, lng = ~longitude, radius = 6, stroke = FALSE, color = "orange", fillOpacity = 0.5)

```

Final Maps

```{r}
library(rmapshaper)
simplified <- rmapshaper::ms_simplify(Mus_MPAs, keep = 0.05)

# Mustelus fasciatus map

Mus_MPAs <- readRDS("Data/Final MPAs/Mus_MPAs.rds")

map01 <-
leaflet(options = leafletOptions(zoomControl = FALSE, minZoom = 5)) %>% 
  addTiles() %>%
  addPolygons(data = simplified, color = "gray", weight = 1, smoothFactor = 0.5,
    opacity = 1.0, fillOpacity = 0.5, fillColor = "tomato") %>%
  addCircleMarkers(data = distributions[["Mustelus fasciatus"]], lat = ~latitude, lng = ~longitude, radius = 5, weight = 3, stroke = TRUE, color = "black", fillOpacity = 0.8) %>%
  addMiniMap(width = 100, height = 100) %>%
  setMaxBounds(lng1 = -60, lat1 = -40, lng2 = -50, lat2 = -29)

# Cephaloscyllium albipinnum map

AUS_MPAs <- readRDS("Data/Final MPAs/Cep_MPAs.rds")

map02 <-
leaflet(options = leafletOptions(zoomControl = FALSE, minZoom = 5)) %>% 
  addTiles() %>%
  addPolygons(data = AUS_MPAs, color = "gray", weight = 1, smoothFactor = 0.5,
    opacity = 1.0, fillOpacity = 0.5, fillColor = "tomato") %>%
  addCircleMarkers(data = distributions[["Cephaloscyllium albipinnum"]], lat = ~latitude, lng = ~longitude, radius = 5, weight = 3, stroke = TRUE, color = "black", fillOpacity = 0.8) %>%
  addMiniMap(position = "bottomleft", width = 100, height = 100) %>%
  setMaxBounds(lng1 = 127, lat1 = -43, lng2 = 152, lat2 = -34)

# Isogomphodon oxyrhynchus map

Iso_MPAs <- readRDS("Data/Final MPAs/Iso_MPAs.rds")

map03 <-
leaflet(options = leafletOptions(zoomControl = FALSE, minZoom = 5)) %>% 
  addTiles() %>%
  addPolygons(data = Iso_MPAs, color = "gray", weight = 1, smoothFactor = 0.5,
    opacity = 1.0, fillOpacity = 0.5, fillColor = "tomato") %>%
  addCircleMarkers(data = distributions[["Isogomphodon oxyrhynchus"]], lat = ~latitude, lng = ~longitude, radius = 5, weight = 3, stroke = TRUE, color = "black", fillOpacity = 0.8) %>%
  addMiniMap(position = "topright", width = 100, height = 100) %>%
  setMaxBounds(lng1 = -63, lat1 = 3, lng2 = -51, lat2 = 12)

# Glyphis garricki map

Gly_MPAs <- readRDS("Data/Final MPAs/Gly_MPAs.rds")

map04 <-
leaflet(options = leafletOptions(zoomControl = FALSE, minZoom = 5)) %>% 
  addTiles() %>%
  addPolygons(data = Gly_MPAs, color = "gray", weight = 1, smoothFactor = 0.5,
    opacity = 1.0, fillOpacity = 0.5, fillColor = "tomato") %>%
  addCircleMarkers(data = distributions[["Glyphis garricki"]], lat = ~latitude, lng = ~longitude, radius = 5, weight = 3, stroke = TRUE, color = "black", fillOpacity = 0.8) %>%
  addMiniMap(position = "topleft", width = 100, height = 100) %>%
  setMaxBounds(lng1 = 122, lat1 = -18, lng2 = 146, lat2 = -6)

```

Heat map test:

```{r}
install.packages("leaflet.extras")
library(leaflet.extras)

write.csv(distributions[[1]], "Data/Shark Data/Mustelus_fasciatus1.csv")

test1 <- read.csv("Data/Shark Data/Mustelus_fasciatus.csv")
test2 <- distributions[[2]]

leaflet() %>%
  addTiles() %>%
  addHeatmap(data = test2, lng = ~longitude, lat = ~latitude, blur = 50, gradient = "YlOrRd")
```

Density ggplot test:

```{r}
install.packages("hexbin")

basemap <- get_map(location = c(left = 127, bottom = -44, right = 152, top = -32.5), maptype = "roadmap", source = "google", force = TRUE)

test <- distributions[[3]] %>% 
  dplyr::rename(lat = latitude, lon = longitude)

shapefile_df <- ggplot2::fortify(AUS_MPAs)

broom::tidy(URY_MPAs)

ggmap(basemap) +
  geom_point(data = test, col = "orange", size = 1) +
  geom_sf(data = URY_MPAs, col = "gray", fill = "tomato")
  stat_density_2d(data = test, aes(fill = ..level.., alpha = ..level..), bins = 50, geom = "polygon")

  # geom_hex(aes(x = lon, y = lat), data = test)

ggplotmap <- ggmap::ggmap(basemap2) +
    coord_equal() +
    labs(x = NULL, y = NULL) +
    theme(axis.text = element_blank())

ggplot() +
  geom_hex(aes(x = lon, y = lat), data = test)

test %>%
  ggplot(aes(x = lon, y = lat)) +
  stat_density_2d(aes(fill = ..level..), geom = "polygon")
  
geom_sf(data = shapefile_df, col = "gray", fill = "tomato")

install.packages("mapproj")
library(ggspatial)

numeric_to_df <- data.frame(URY_MPAs$geometry[[1]][[1]][[1]])

# ggplot() +
ggplotmap +
  ggspatial::geom_spatial_polygon(aes(x = lon, y = lat), data = numeric_to_df, fill = "tomato", colour = "black") +
  geom_point(data = test, col = "orange", size = 1)
  theme_void() +
  coord_map()
```

Converting back to shp file

```{r}
URY_MPAs <- readRDS("Data/MPAs by country/URY_MPAs.rds")

rgdal::writeOGR(URY_MPAs, "Data/MPAs by country", "URY_MPAs_shp", driver = "ESRI Shapefile")
```

